---
title: "Soil Effects on Presence/Absence Analysis"
output: html_notebook
---
```{r}
data2009 <-read.csv("2009.Data.Entry.Copy.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)
str(data2009)
data2018 <-read.csv("2018.Data.Entry.Copy.csv", header = TRUE, sep = ",",stringsAsFactors = FALSE)
str(data2018)

```
pulled in both data sets of data from csv files 
```{r}
d18<-data2018$SiteName[data2018$SiteName=="Naimen"]<-"Naiman"

```
renamed the variable Naimen to Naiman to match 2009 data and created a new dataframe with manipulated data
```{r}
d09<-rename(data2009,Transect = SiteCode)
```
renamed sitecode to transect in 2009 data to match 2018 and created a new dataframe with manipulated 
```{r}
install.packages("reshape")
d09pc <- cast(d09,SiteName + PlotNumber ~ Species_Code, value='Cover', FUN=mean)

```
installed reshape package and casted data by species type and site code. Used cover as value to track basic presence/absence for 2009 data
```{r}
library(dplyr)
d18<-rename(data2018,Species_Code = SpeciesCode)
```
reloaded dplyr pakage, and renamed SpeciesCode to match d09 
```{r}
d18pc <- cast(d18,SiteName + PlotNumber ~ Species_Code, value='Cover', FUN=mean)
```
performed same cast function for d18 to obtain presence absence 
```{r}
library(dplyr)
d18<-rename(data2018,Species_Code = SpeciesCode)
d18pc <- cast(d18,SiteName + PlotNumber ~ Species_Code, value='Cover', FUN=mean)
```
load dplyr package, created new data frame for manipulation of data called d18. Casted d18 into d18pc for presence/absence. 
```{r}
statsd09 <- d09 %>% group_by(SiteName) %>% group_by(Species_Code) %>% summarize(avg = mean(Cover))
statsd18 <- d18 %>% group_by(SiteName) %>% group_by(Species_Code) %>% summarize(avg = mean(Cover))
View(statsd09)
View(statsd18)
```
Created new data frame that contains the summary cover statistics of mean cover for each species within 2009 data 
repeated the same steps for 2018 data. Used View function to confirm data has been subsetted in the way that I wanted
```{r}
summd09pa<- cast(statsd09,~Species_Code, value='avg', FUN=mean)
summd18pa<- cast(statsd09,~Species_Code, value='avg', FUN=mean)
```
cast summary data to view the species cover by average 
```{r}
d09pc <- as.data.frame(d09pc)
d18pc <- as.data.frame(d18pc)
```
changing both casts from 2009 and 2018 to data frames so that NA values can be converted into 0
```{r}
d09pc[is.na(d09pc)] <- 0
d18pc[is.na(d18pc)] <- 0
```
changed NA to 0
```{r}
d09pc.matrix <- as.matrix.cast_df(d09pc)
d18pc.matrix <- as.matrix.cast_df(d18pc)
```
casted to a matrix 
```{r}
naiman09.avg <- d09 %>% filter(SiteName == "Naiman") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
grazing09.avg <- d09 %>% filter(SiteName == "Grazing") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
```
calculated mean cover by site, grouped by species for 2009 data. 
```{r}
naiman18.avg <- d18 %>% filter(SiteName == "Naiman") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
control18.avg <- d18 %>% filter(SiteName == "Control") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
grazing18.avg <- d18 %>% filter(SiteName == "Grazing") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
heavy18.avg <- d18 %>% filter(SiteName == "Heavy") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
light18.avg <- d18 %>% filter(SiteName == "Light") %>% group_by(Species_Code) %>% summarise(avg = mean(Cover))
```
ran the same function for 2018 data 
```{r}
med.site.d18 <- d18 %>% group_by(SiteName) %>% summarize(median.of.all.sites = median(Cover))
med.site.d09 <- d09 %>% group_by(SiteName) %>% summarize(median.of.all.sites = median(Cover))
```

found median cover and std for species by sites for 2009 and 2018 data and assigned to their own data frame 
```{r}
naiman09.med <- d09 %>% filter(SiteName == "Naiman") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
grazing09.med <- d09 %>% filter(SiteName == "Grazing") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
naiman09.sd <- d09 %>% filter(SiteName == "Naiman") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
grazing09.sd <- d09 %>% filter(SiteName == "Grazing") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
naiman18.med <-d18 %>% filter(SiteName == "Naiman") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
control18.med <- d18 %>% filter(SiteName == "Control") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
grazing18.med <-d18 %>% filter(SiteName == "Grazing") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
heavy18.med <-d18 %>% filter(SiteName == "Heavy") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
light18.med <-d18 %>% filter(SiteName == "Light") %>% group_by(Species_Code) %>% summarise(med = median(Cover))
naiman18.sd <-d18 %>% filter(SiteName == "Naiman") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
control18.sd <-d18 %>% filter(SiteName == "Control") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
grazing18.sd <-d18 %>% filter(SiteName == "Grazing") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
heavy18.sd <-d18 %>% filter(SiteName == "Heavy") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
light18.sd <-d18 %>% filter(SiteName == "Light") %>% group_by(Species_Code) %>% summarise(std = sd(Cover))
```
Attempted a table join of Naiman 2009 summary statistics. 
```{r}
naiman09.allstats <- left_join(naiman09.avg,naiman09.med, by = "Species_Code")
View(naiman09.allstats)
naiman09.summ.stat <- left_join(naiman09.allstats,naiman09.sd, by = "Species_Code")
```
found I could not join more than 2 tables at atime, so I ran separate joins for STD, however found that by piping I could complete the process faster, with less code. So I completed the join of the data tables of summary statistics for remaining sites in both years 2009 and 2018 data as grouped species 
```{r}
naiman18.summ.stats <- left_join(naiman18.avg,naiman18.med, by='Species_Code') %>% left_join(.,naiman18.sd, by='Species_Code')
grazing18.summ.stats <- left_join(grazing18.avg,grazing18.med, by='Species_Code') %>% left_join(.,grazing18.sd, by='Species_Code')
control18.summ.stats <- left_join(control18.avg,control18.med, by='Species_Code') %>% left_join(.,control18.sd, by='Species_Code')
heavy18.summ.stats <- left_join(heavy18.avg,heavy18.med, by='Species_Code') %>% left_join(.,heavy18.sd, by='Species_Code')
light18.summ.stats <- left_join(light18.avg,light18.med, by='Species_Code') %>% left_join(.,light18.sd, by='Species_Code')
grazing09.summ.stats <- left_join(grazing09.avg,grazing09.med, by='Species_Code') %>% left_join(.,grazing09.sd, by='Species_Code')
```
re-did table join for Naiman 2009 so that naming convention and process matched the rest of the data 
```
```{r}
naiman09.summ.stats <- left_join(naiman09.avg,naiman09.med, by='Species_Code') %>% left_join(.,naiman09.sd, by='Species_Code')
```


